---
title: "Strawberries"
author: "Haoran Su"
date: "10/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(citation)
opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```
#Strawberries
##Background
The data were collected from the USDA database selector: <a href="https://quickstats.nass.usda.gov">https://quickstats.nass.usda.gov</a>

The data were <a href="https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1">stored online</a> and then downloaded as a CSV file.

In this assignment, after cleaning and reorganizing data, we select the strawberries from three kinds of berries and saperate columns to leave the required data.
Then we make boxplots for strawberries of years and states. It can also be visualized on the shiny app.

## Data cleaning and reorganizing
```{r,echo=FALSE}

## read the data

ag_data <- read_csv("berries.csv", col_names = TRUE)

## look at number of unique values in each column
ag_data %>% summarize_all(n_distinct) -> aa


## make a list of the columns with only one unique value
bb <- which(aa[1,]==1)

## list the 1-unique value column names 
cn <- colnames(ag_data)[bb]

```

Data selected from the NASS database often has columns without any data or with a single repeated Values.  The berries data had only 8 out of 21 columns containing meaningful data.

```{r,echo=FALSE}
## remove the 1-unique columns from the dataset
ag_data %<>% select(-all_of(bb))
aa %<>% select(-all_of(bb)) 

## State name and the State ANSI code are (sort of) redundant
## Just keep the name
ag_data %<>% select(-4)
aa %<>% select(-4) 

kable(head(ag_data)) %>%
  kable_styling(font_size=12)

## Display the head of ag_data 
kable(head(ag_data)) %>%
  kable_styling(font_size=12)

## 3 kinds of berries
sberry <- unique(ag_data$Commodity)
```

This table contains information about berries: blueberries, raspberries, and strawberries. We only choose Strawberries for analyzing and data need further cleaning and organizing. Only YEAR" time period will be considered.
Also, we separate the `Data Item`, `Domain`, `Domain Category` columns for further cleaning process and remove the redundant columns. 

```{r,echo=FALSE}
# separate the "strawberries"
sberry <- ag_data %>% filter((Commodity=="STRAWBERRIES") & (Period=="YEAR"))
sberry %<>% select(-c(Period, Commodity)) 

# separate "Data Item"
sberry %<>% separate(`Data Item`, c("B","type", "meas", "what"), sep = ",") 
sberry  <- select(sberry, -B)
sberry %<>% separate(type,c("s1", "type", "s2", "lab1", "lab2"), " ")
sberry %<>% select(-c(s1,s2)) 
#remove all NA in dataset to " "
sberry[is.na(sberry)] <- " " 

# separate "Domain" and "Domain Category"
sberry %<>% separate(Domain, c("D_left", "D_right"), sep = ", ")
sberry %<>% separate(`Domain Category`, c("DC_left", "DC_right"), sep = ", ")
sberry %<>% separate(DC_left, c("DC_left_l", "DC_left_r"), sep = ", ")
sberry %<>% separate(DC_right, c("DC_right_l", "DC_right_r"), sep = ", ") 
#remove all NA in dataset to " "
sberry[is.na(sberry)] <- " "

sberry %<>% select(-DC_left_l) 
sberry %<>% select(-DC_right_l)
sberry %<>% mutate(label = paste(lab1,lab2)) 
sberry %<>% mutate(D_left = "CHEMICAL", D_left = "") 
sberry %<>% mutate(Chemical=paste(D_left, D_right)) 
sberry %<>% select(-c(D_left, D_right)) 
sberry %<>% select(State, Year, type, what, meas, label, DC_left_r, DC_right_r, Chemical, Value )
```


Within the "what" and "meas" columns, there are data begins with "MEASURED IN".We want to put them all in the same column.

```{r,echo=FALSE }
## So, we will separate them from their current column and put them into two columns -- then we will test to make sure there aren't any overlaps and then merge the two columns
#make a function f1
f1 <- function(a,b){
  if(a){
    return(b)
  }else{
      return("")
  }
}

### index of cells to be isolated
index_meas <- str_detect(sberry$meas, "MEASURED IN")

### create a column of the isolated cells
sberry %<>% mutate(m_in_1 = unlist(map2(index_meas, sberry$meas, f1))) 

###  eliminate the isolated cells from the original column
sberry %<>% mutate(meas = str_replace(sberry$meas, "MEASURED IN.*$", ""))

## index of cells to be isolated
index_what <- str_detect(sberry$what, "MEASURED IN")

### create a column of the isolated cells
sberry %<>% mutate(m_in_2 = unlist(map2(index_what, sberry$what, f1))) 

###  eliminate the isolated cells from the original column
sberry %<>% mutate(what = str_replace(sberry$what, "MEASURED IN.*$", ""))

### Check for overlaps
sberry %<>% mutate(units = str_trim(paste(m_in_1, m_in_2))) 
sberry$units %>% unique()

```


Finally, we rename the columns of the dataset and select those that are required in this analysis.

```{r,echo=FALSE}
sberry %<>% rename(c(Avg = what, Marketing = meas, Harvest = label, 
                               Chem_family = DC_left_r, Materials = DC_right_r, 
                               Measures = units))

str_trim(paste(sberry$Marketing, sberry$Harvest)) %>% unique
sberry %<>% mutate(production = str_trim(paste(Marketing, Harvest)))
sberry %<>% mutate(Chemical = str_trim(paste(Chem_family, Chemical)))

sberry %<>% select(Year, State, type, production, Avg, Measures, Materials, Chemical, Value)

#display the headings of current data "sberry"
colnames(sberry)
```
After cleaning and reorganizing the columns, current headings of data "sberry" is as above. As the chemicals of strawberries is the factor we want to analysize, we do further filtering for the data.

```{r,echo=FALSE}
unfood <- sberry %<>% filter(production=="APPLICATIONS") 

unfood %<>% filter(Value != "(D)")

unfood %<>% filter(Value !=  "(NA)")

unfood %<>% filter(Measures == "MEASURED IN LB / ACRE / APPLICATION")

df<- unfood %>%  select(Year, State, Chemical, Value, type, production)

df$Value = as.numeric(df$Value)

```

1. We make a box plot of chemical values being applied to strawberries according to different states. 

```{r,echo=FALSE}
# box plot of different states and value
bp1 <- ggplot(df, aes(x = State, y = Value))
bp1 <- bp1 + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 13, face = "bold")) +
  labs(x = "Different states")
bp1
```
From the plot, we find many outliers influencing the presentation so we remove the outliners and make another box plot.

```{r,echo=FALSE}
bp2 <- ggplot(df, aes(x = State, y = Value))
bp2 <- bp2 + geom_boxplot(outlier.colour = NA) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 13, face = "bold")) +
  coord_cartesian(ylim = c(0, 8e+6)) +
  labs(x = "Different states")
bp2
```
From the plot, we can see that in most states, the chemical values of straberries are quite small. Only in Ohio, Pennsylvania, Wisconsin and other states, the chemical value is high.


2. Then we make a box plot of chemical values being applied to strawberries according to years. 

```{r}
# boxplot of different years
bp3 <- ggplot(df, aes(x = factor(Year), y = Value))
bp3 <- bp3 + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 13, face = "bold")) +
  labs(x = "Different years")
bp3
```
Similarly, the box plot of years also have many outliers so we remove them and get another bow plot.

```{r}
# excluding outliers
bp4 <- ggplot(df, aes(x = factor(Year), y = Value))
bp4 <- bp4 + geom_boxplot(outlier.colour = NA) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 13, face = "bold")) +
  coord_cartesian(ylim = c(0, 6e+6)) +
  labs(x = "Different years")
bp4
```
Within the years 2015-2019, chemical values of strawberries is quite low in 2016, 2018 and 2019. It has the highest values in 2015 and year 2017 gets the second rank.



## Citation
*
David Morison (2017) Exploratory data analysis into the relationship between different types of crime in London. Available at: https://towardsdatascience.com/exploratory-data-analysis-into-the-relationship-between-different-types-of-crime-in-london-20c328e193ff (Accessed: 15 Oct 2020).
*
United States Department of Agriculture National Agricultural Statistics Service (n.p.) Quick Stats. Available at: https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1 (Accessed: 15 Oct 2020).

```{r}
package2zenodo("knitr")
package2zenodo("tidyverse")
package2zenodo("kableExtra")
package2zenodo("citation")
```